Destruction of metrizability in generalized inverse limits

### 2015-09-17 articles/destroying-metrizability-in-generalized-inverse-limits/article.tex
mention nonempty compactness, connectedness for kappa=omega. elaborate on corson compactness

@@

Let us say a subset $S$ of $[0,1]^2$ satisfies condition $\Gamma$ if, for some $x, y \in [0,1]$ with $x \ne y$, $\<x,x\> \in S$, $\<y,y\> \in S$, and either $\<x,y\>$ or $\<y,x\> \in S$. Let $\Delta = \{\<a,a\> \ | \ a \in [0,1]\}$, and notice that $[0,1]^2 \setminus \Delta$ is the union of two disjoint open sets.

@@

A \textit{Corson compact} space is a compact space which embeds in a $\Sigma$-product of real lines (see \cite{alster}). All compact metric spaces are embeddable in $\mathbb R^\omega$ and therefore Corson compact.

@@

\begin{proof}
If $f = \iota$, then of course the inverse limit is homeomorphic to $[0,1]$. So, suppose $f \ne \iota$. If $\kappa$ is countable, then because the inverse limit is a subspace of $[0,1]^{\kappa}$, it is metric. However, if $\kappa$ is uncountable, then by Theorem \ref{main theorem}, the inverse limit contains a copy of the non-Corson compact space $\kappa + 1$.
\end{proof}

@@

The discussion of this example in \cite{varagona} implies that this inverse limit is homeomorphic to $Y = (\kappa \times [0,1)) \cup \{\< \kappa, 0 \>\}$ (with the lexicographic order topology) whenever $\kappa$ is a nonzero limit ordinal. If $0 < \kappa < \omega_1$, $Y$ is a metric arc. However, when $\kappa = \omega_1$, $Y$ is the compactified long line, a non-metric space.
\

@@

>>>>

Ingram and Mahavier showed in \cite{i m paper} that $\varprojlim \{X_{\alpha}, f_{\alpha \beta}, \omega\}$ is non-empty and compact. They also found that
if $f_{\alpha \beta}$ is always continuum-valued, then $\varprojlim \{X_{\alpha}, f_{\alpha \beta}, \omega\}$ is also connected, and therefore a Hausdorff continuum. Greenwood and Kennedy showed in \cite{greenwood kennedy} that there is a non-continuum-valued u.s.c. surjective idempotent bonding map $f$ for which $\varprojlim \{X_{\alpha}, f_{\alpha \beta}, \omega\}$ is disconnected.

@@

Let us say a subset $S$ of $[0,1]^2$ satisfies condition $\Gamma$ if, for some $x, y \in [0,1]$ with $x \ne y$, $\<x,x\> \in S$, $\<y,y\> \in S$, and either $\<x,y\>$ or $\<y,x\> \in S$; see Figure \ref{conditionGamma}. Let $\Delta = \{\<a,a\> \ | \ a \in [0,1]\}$, and notice that $[0,1]^2 \setminus \Delta$ is the union of two disjoint open sets.

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{idempotent_usc_4.pdf}
\end{center}
\caption{Non-trivial u.s.c. surjections satisfying condition $\Gamma$}
\label{conditionGamma}
\end{figure}

@@

Note that when $\kappa\geq\omega_1$, $\omega_1\in\kappa+1$ is a point of non-first-countability; therefore any space containing $\kappa+1$ as a subspace is nonmetrizable.

The presence of $\kappa+1$ prevents more than just metrizability, however. Let $\Sigma\mathbb R^\kappa\subseteq \mathbb R^\kappa$ denote the \textit{Sigma-product} of $\kappa$ copies of $\mathbb R$: the subspace for which each sequence has only countably many coordinates of nonzero value. A \textit{Corson compact} space is a compact space which embeds in $\Sigma\mathbb R^\kappa$ for some $\kappa$ (see \cite{alster}).

All compact metric spaces are embeddable in $\mathbb R^\omega=\Sigma\mathbb R^\omega$ and therefore Corson compact. In addition, every closed subspace of a Corson compact is Corson compact. While $\Sigma\mathbb R^\kappa$ is not first-countable for $\kappa\geq\omega_1$, any subspace holds the weaker notion of being a $W$-space as defined by Gruenhage in \cite{gruenhage}. However, $\omega_1+1$ does not, as every countable sequence of ordinals less than $\omega_1$ is bounded above and therefore cannot converge to $\omega_1$ itself; therefore any space containing $\kappa+1$ for $\kappa\geq\omega_1$ cannot even be Corson compact.

@@

\begin{proof}
If $f = \iota$, then of course the inverse limit is homeomorphic to $[0,1]$. So, suppose $f \ne \iota$. If $\kappa$ is countable, then because the inverse limit is a subspace of $[0,1]^{\kappa}$, it is metrizable. However, if $\kappa$ is uncountable, then by Theorem \ref{main theorem}, the inverse limit contains a copy of $\kappa + 1$, preventing metrizability and Corson compactness.
\end{proof}

@@

The discussion of this example in \cite{varagona} implies that this inverse limit is homeomorphic to $Y = (\kappa \times [0,1)) \cup \{\< \kappa, 0 \>\}$ (with the lexicographic order topology) whenever $\kappa$ is a nonzero limit ordinal. If $0 < \kappa < \omega_1$, $Y$ is a metric arc. However, when $\kappa = \omega_1$, $Y$ is the compactified long line, a non-metric space.

@@

\bibitem{greenwood kennedy} Greenwood, Sina and Kennedy, Judy, {\it Connected generalized inverse limits}, Topology and its Applications,
159 (2012), no. 1, 57–68.

\bibitem{gruenhage} G. Gruenhage, {\it Infinite games and generalizations of first-countable spaces}, Gen. Topol. Appl., 6 (1976), pp. 339–352
